Homework 4 - Stat 495

### 2022-10-09 Midterm Materials/MidtermPracticeQuestions.Rmd
Hwk 4 before changing

## Add 2

@@

## Add 3

@@

## Add 4

@@

## Add 5

@@

## Add 6

@@

(a) Use OLS to predict *prog* using *diabetes2* and verify you obtain the results in Table 7.3.
(b) Use ridge regression with lambda = 0.1 and verify you obtain the results in Table 7.3.
(c) Explain what cross-validation is (generally), and how it can be used to help choose a lambda in ridge regression.
(d) Use CV with ridge regression to select your lambda running the default grid through glmnet (i.e. don't specify a grid). What lambda is selected?
(e) Implement LASSO with CV to select your lambda running the default grid through glmnet. What lambda is selected? 
(f) Implement a regression tree to predict *prog*. What cp corresponds to the default tree? 
(g) Explain why we don't need a *glm* here to predict *prog*. 
(h) Which model of these do you prefer? How do the fits differ? (Hint: Besides comparing coefficients/involved variables, you can compare MSEs.)

## Add 7

@@

(a) Why does it not make sense to use OLS to predict *spam*?
(b) Use logistic regression to predict *spam*. Verify you obtain the results in Table 8.3.
(c) Obtain a confusion matrix for the logistic regression. How well is the model doing?
(d) Use a classification tree to predict *spam*. Attempt to obtain the tree in Figure 8.7. Note that you may need to prune or adjust control parameters to obtain this tree. If you cannot obtain the tree, obtain one you want to work with for part e below.
(e) Obtain a confusion matrix for your classification tree. How do your error rates compare to those reported in Figure 8.7? 
(f) Which method do you prefer for predicting *spam*?

## CASI 8.6

@@

## Add 8

@@

(a) Create an appropriate training/test split using a 75/25 percent split. 
(b) Perform best subsets and choose a model based on an appropriate descriptive statistic. Compute the test MSE.  
(c) Perform forward selection and choose a model based on an appropriate descriptive statistic. Compute the test MSE.  
(d) Perform the lasso and choose a model based on an appropriate descriptive statistic. Compute the test MSE.  
(e) Fit an elastic-net penalty with alpha = 0.5, and choose a model based on an appropriate descriptive statistic. Compute the test MSE.  
(f) How do your models compare? Which model do you prefer? Why? (Should be able to compare coefficients.)

## Add 9

>>>>

\newpage
## Add 2

@@

\newpage
## Add 3

@@

\newpage
## Add 4

@@

\newpage
## Add 5

@@

\newpage
## Add 6

@@

(a) Use OLS to predict *prog* using *diabetes2* and verify you obtain the results in Table 7.3.

```{r}
mod1 <- lm(prog ~ ., data = diabetes2)
msummary(mod1)
round(coef(mod1), 2)
```

(b) Use ridge regression with lambda = 0.1 and verify you obtain the results in Table 7.3.
(c) Explain what cross-validation is (generally), and how it can be used to help choose a lambda in ridge regression.
(d) Use CV with ridge regression to select your lambda running the default grid through glmnet (i.e. don't specify a grid). What lambda is selected?

```{r}
x2 <- model.matrix(prog ~ ., diabetes)[, -1]
y2 <- diabetes %>% 
  select(prog) %>%
  unlist() %>%
  as.numeric()

set.seed(1)
cv.out <- cv.glmnet(x2, y2, alpha = 0)
bestlam <- cv.out$lambda.min
bestlam

ridge_mod <- glmnet(x2,y2, alpha = 0)
round(coefficients(ridge_mod, s = bestlam), 3)
```

(e) Implement LASSO with CV to select your lambda running the default grid through glmnet. What lambda is selected?

```{r}
set.seed(1)
cv.out1 <- cv.glmnet(x2, y2, alpha = 1)
bestlam1 <- cv.out1$lambda.min
bestlam1

lasso_mod <- glmnet(x2,y2, alpha = 1)
round(coefficients(lasso_mod, s = bestlam1), 3)
```

(f) Implement a regression tree to predict *prog*. What cp corresponds to the default tree? 
```{r}
set.seed(1)
prog.rpart <- rpart(prog ~ ., data = diabetes2, method = "anova")
printcp(prog.rpart)
```
(g) Explain why we don't need a *glm* here to predict *prog*. 
`prog` is a continuous, numeric variable so OLS is fine. GLMs are used when there is logistic, poisson for counts, etc.

(h) Which model of these do you prefer? How do the fits differ? (Hint: Besides comparing coefficients/involved variables, you can compare MSEs.)

```{r}
# for OLS
54.16^2

# for ridge
ridge_pred <- predict(ridge_mod, s = bestlam, newx = x2)
mean((ridge_pred - y2)^2)

# for lasso
lasso_pred <- predict(lasso_mod, s = bestlam1, newx = x2)
mean((lasso_pred - y2)^2)

# for tree
fittedtree <- predict(prog.rpart)
mean((fittedtree - diabetes2$prog)^2)
```

\newpage
## Add 7

@@

(a) Why does it not make sense to use OLS to predict *spam*?
It doesn't make sense to use OLS to predict spam because `spam` is a 0/1 variable and OLS doesn't have a way to constrain the response variable to be 0/1.
(b) Use logistic regression to predict *spam*. Verify you obtain the results in Table 8.3.
```{r}
spamdata <- select(spam, -spam, -testid)
spamdata <- data.frame(scale(spamdata))
spamdata <- mutate(spamdata, spam = spam$spam)
```

```{r}
loglm <- glm(spam ~ ., data = spamdata, family = "binomial")
summary(loglm)
```

```{r}
tally(~ spam, data = spamdata)
1813/4601
```

```{r}
logaug <- loglm %>%
  broom::augment(type.predict = "response")
logaug <- mutate(logaug, binprediction = round(.fitted,0))
tally(~binprediction, data = logaug)
```
```{r}
with(logaug, table(spam, binprediction))
```

(c) Obtain a confusion matrix for the logistic regression. How well is the model doing?
(d) Use a classification tree to predict *spam*. Attempt to obtain the tree in Figure 8.7. Note that you may need to prune or adjust control parameters to obtain this tree. If you cannot obtain the tree, obtain one you want to work with for part e below.
```{r}
set.seed(495)
spam.rpart <- rpart(factor(spam) ~ ., method = "class", data = spamdata)
printcp(spam.rpart)
#rplot.plot(spam.rpart)
```

(e) Obtain a confusion matrix for your classification tree. How do your error rates compare to those reported in Figure 8.7?

```{r}
spampredict <- predict(spam.rpart, type = "class")
table(spampredict, spamdata$spam)
```

(f) Which method do you prefer for predicting *spam*?
The tree has a little worse performance, but is much easier to understand so we would probably use thr tree.
\newpage
## CASI 8.6

@@

```{r}
galaxy2 <- mutate(galaxy, m = 18:1)

galaxydata <- galaxy2 %>%
  tidyr::pivot_longer(-m, names_to = "log_r", values_to = "count")

galaxydata <- mutate(galaxydata, r = c(rep(1:15, 18)))
```

```{r}
myglm <- glm(count ~ r*m + I(r^2) + I(m^2), data = galaxydata, family = poisson)
lmtest::lrtest(myglm)
```

\newpage
## Add 8

@@

(a) Create an appropriate training/test split using a 75/25 percent split.

(b) Perform best subsets and choose a model based on an appropriate descriptive statistic. Compute the test MSE.  
(c) Perform forward selection and choose a model based on an appropriate descriptive statistic. Compute the test MSE.  
(d) Perform the lasso and choose a model based on an appropriate descriptive statistic. Compute the test MSE.  
(e) Fit an elastic-net penalty with alpha = 0.5, and choose a model based on an appropriate descriptive statistic. Compute the test MSE.  
(f) How do your models compare? Which model do you prefer? Why? (Should be able to compare coefficients.)

\newpage
## Add 9

### 2022-10-09 homework/hmk4/hmk4Stat495F22.Rmd
Hwk 4 before changing

# we are only looking at prices greater than 500,00 and change date to a 
# date format
kchouse <- kchouse %>%
  filter(price > 500000) %>%
  mutate(date = lubridate::mdy(date))

# create a histogram
gf_dhistogram(~ price, data = kchouse) %>%
  gf_labs(title = "Histogram of Prices Greater than $500,000")

gf_boxplot(~ price, data = kchouse)%>%
  gf_labs(title = "Boxplot of Prices Greater than $500,000")

@@

The distribution of `price` is right-skewed with a median of \$690,000 and an IQR is \$294,000. There appears to be many outliers on the high end for `price`. We are only interested in predicting when `price` is greater than \$500,000, `price` is bounded on the lower end at \$500,000. Since the distribution is right-skewed, a log-transformation could be useful.

>>>>

# change date to a 
# date format

kchouse <- kchouse %>%
  mutate(date = lubridate::mdy(date))

# create a histogram
gf_dhistogram(~ price, data = kchouse) %>%
  gf_labs(title = "Histogram of Prices")

gf_boxplot(~ price, data = kchouse)%>%
  gf_labs(title = "Boxplot of Prices")

@@

```{r}
645000-322000
```

The distribution of `price` is right-skewed with a median of \$450,000 and an IQR is \$323,000. There appears to be many outliers on the high end for `price`. The minimum of `price` is \$78,000 and the maximum is \$7,700,000. Since the distribution is right-skewed, a log-transformation could be useful.
